Tropical Andes plant Lookup Table

### 2023-11-15 code/L1/L1_P2_TropicalAndes_taxcleaning.Rmd
add nrow check code and update column names

---
title: "Tropical Andes plant Lookup Table"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-07-25"
output: html_document
---

@@

# Use bdc to check scientific names 
```{r}
name_check_results <- bdc_clean_names(sci_names = all_species, save_outputs = FALSE)
```
```{r}
all_species_cleaned <- name_check_results$names_clean
```

@@

# Subset TNRS results to species with matches, conflicts, and disparties
```{r}
unique(all_species_hamonized_results$Taxonomic_status)
```

@@

# Disparties
all_species_no.opinion <- subset(all_species_hamonized_results, Taxonomic_status == "No opinion")
nrow(all_species_no.opinion)
all_species_illegitimate <- subset(all_species_hamonized_results, Taxonomic_status == "Illegitimate")
nrow(all_species_illegitimate)
all_species_invalid <- subset(all_species_hamonized_results, Taxonomic_status == "Invalid")
nrow(all_species_invalid)
all_species_other <- subset(all_species_hamonized_results, Taxonomic_status == "")
nrow(all_species_other)
```
# Create species list of names with disparities
```{r}
all_species_disparities <- all_species_hamonized_results %>%
                            filter(Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", "")) %>%
                            select(Name_submitted)
```

# Use bdc to harmonize disparities
## Using ITIS database
```{r}
all_species_disparities_col_results <- bdc_query_names_taxadb(sci_name = all_species_disparities$Name_submitted, replace_synonyms = TRUE,suggest_names = TRUE, db = "col", rank_name = "Plantae", rank = "kingdom")
```

## Subset COL results to accepted, synonym, and NA

```{r}
unique(all_species_disparities_col_results$taxonomicStatus)
```

@@

# Create lookup table
## From TNRS
```{r}
species_matches <- all_species_matches[ , c("Name_submitted", "Name_matched", "Source")]
species_corrected <- all_species_conflicts[ , c("Name_submitted", "Name_matched", "Source")]
```

@@

## From bdc using COL
```{r}
species_col_accepted <- all_species_col_accepted[ , c("original_search", "suggested_name")]
species_col_accepted$source <- c("col")
species_col_synonmyn <- all_species_col_synonym[ , c("original_search", "suggested_name")]
species_col_synonmyn$source <- c("col")
```

```{r}
#update col names
colnames(species_col_accepted) <- c("Name_submitted", "Name_matched", "Source")
colnames(species_col_synonmyn) <- c("Name_submitted", "Name_matched", "Source")
```

@@

# Update data files species names using lookup table
## GBIF
```{r}
# filter data to species in lookup_table$Names_submitted
GBIF_occ_subset <- GBIF_occ %>%
  filter(species %in% lookup_table$Name_submitted)
nrow(GBIF_occ_subset)
# replace species with lookup_table$Name_matched
GBIF_occ_subset_harmonized <- GBIF_occ_subset %>% mutate(species = replace(species, species %in% lookup_table$Name_submitted, lookup_table$Name_matched))
nrow(GBIF_occ_subset_harmonized)
```

# Summary
```{r}

@@

>>>>

---
title: "Tropical Andes plant Lookup Table"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-07-25"
output: 
  html_document: 
    toc: yes
    df_print: tibble
---

@@

# Use bdc to check scientific names 
```{r}
name_check_results <- bdc_clean_names(sci_names = all_species, save_outputs = FALSE)
```
```{r}
all_species_cleaned <- name_check_results$names_clean
length(all_species_cleaned)
```

@@

# Subset TNRS results to species with matches, conflicts, and disparties
```{r}
nrow(all_species_hamonized_results)
unique(all_species_hamonized_results$Taxonomic_status)
```

@@

# Disparties
all_species_no.opinion <- subset(all_species_hamonized_results, Taxonomic_status == "No opinion")
nrow(all_species_no.opinion)
all_species_illegitimate <- subset(all_species_hamonized_results, Taxonomic_status == "Illegitimate")
nrow(all_species_illegitimate)
all_species_invalid <- subset(all_species_hamonized_results, Taxonomic_status == "Invalid")
nrow(all_species_invalid)
all_species_other <- subset(all_species_hamonized_results, Taxonomic_status == "")
nrow(all_species_other)
```
# Create species list of names with disparities
```{r}
all_species_disparities <- all_species_hamonized_results %>%
                            filter(Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", "")) %>%
                            select(Name_submitted)
nrow(all_species_disparities)
```

# Use bdc to harmonize disparities
## Using COL database
```{r}
all_species_disparities_col_results <- bdc_query_names_taxadb(sci_name = all_species_disparities$Name_submitted, replace_synonyms = TRUE,suggest_names = TRUE, db = "col", rank_name = "Plantae", rank = "kingdom")
```

## Subset COL results to accepted, synonym, and NA
```{r}
unique(all_species_disparities_col_results$taxonomicStatus)
```

@@

# Create lookup table
## From TNRS
```{r}
species_matches <- all_species_matches[ , c("Name_submitted", "Accepted_species", "Source")]
species_corrected <- all_species_conflicts[ , c("Name_submitted", "Accepted_species", "Source")]
```

@@

## From bdc using COL
```{r}
species_col_accepted <- all_species_col_accepted[ , c("original_search", "scientificName")]
species_col_accepted$source <- c("col")
species_col_synonmyn <- all_species_col_synonym[ , c("original_search", "scientificName")]
species_col_synonmyn$source <- c("col")
```

```{r}
#update col names
colnames(species_col_accepted) <- c("Name_submitted", "Accepted_species", "Source")
colnames(species_col_synonmyn) <- c("Name_submitted", "Accepted_species", "Source")
```

@@

# Update data files species names using lookup table
## GBIF
```{r}
# filter data to species in lookup_table$Names_submitted
GBIF_occ_subset <- GBIF_occ %>%
  filter(species %in% lookup_table$Name_submitted)
nrow(GBIF_occ_subset)
# add accepted species column from lookup table based on species column
GBIF_occ_subset_harmonized <- merge(GBIF_occ_subset, lookup_table, by.x = "species", by.y = "Name_submitted")
nrow(GBIF_occ_subset_harmonized)
```

## TRY
```{r}

@@

## BIEN
```{r}

## GIFT
```{r}

# Summary
```{r}
source("~/GitHub/neotropical_plants/functions/data_summary.R")
```

```{r}
data_summary(GBIF_occ_subset_harmonized, GBIF_occ_subset_harmonized$species, GBIF_occ_subset_harmonized$genus, GBIF_occ_subset_harmonized$family)
```
